Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-04 WSSSPE2_report/WSSSPE2_review.tex
Section 7: minor changes. Not done yet. Commit before taking a short break.

Group discussion started with a position statement by each person surrounding
what they had learned from their software development experiences and how those
lessons might be translated into actionable outcomes. Participants came from a
range of backgrounds, and represented multiple software and training
initiatives, including Software Carpentry, Data Carpentry, Open Science for
Synthesis, the Community Surface Dynamics Modeling Systems group, ROpenSci,
DataONE, the HDF Group, and others. Some of these software experiences were
focused on development of new products for use in the sciences (e.g., ROpenSci,
CSDMS), and these recounted the difficulties of engaging with disciplinary
scientists in writing software. Software was clearly utilized broadly across the
various science disciplines represented, and it was developed within those
disciplines as well. Some researchers created software for statistical analysis
and modeling, while others used it to control instrumentation, collate data
across networks, collect information from respondents, and many other uses.
There was broad consensus that, within the disciplines represented, formal
training in any type of software development or engineering was rare among the
practitioners; most are self-taught, and develop software to get another job
done. Any ancillary utility of the software outside of the specific science
target was generally unplanned and few researchers would want to invest more
time to make their own software more re-usable.

@@

Three main actions were identified by Subgroup A that would be of interest to
participants and that would benefit the research community. These focused on the
desire to amplify the current community efforts in training and software
engineering by connecting the current training initiatives (e.g., software/data
carpentry, rOpenSci, OSS) that are of different durations at appropriate stages.
Generally it was felt that a modicum of interaction between short term workshops
(SwC, DC), medium term trainings (e.g., OSS), and longer-term courses (e.g.,
BIDS Data Science) would benefit from coordinating curriculum, discussing and
aligning prerequisites, and coordinating timing of courses. A training
coordination effort would go a long ways towards amplifying the value of the
individual efforts and make them all more effective.

@@

Subgroup B then focused on actions it could take. It discussed development of a
white paper that describes a matrix approach to training (multi-day, multi-week,
semester). The white paper might include a survey of existing techniques. There
are many, some dating back many years, for example the Interuniversity
Consortium for Political and Social Research (ICPSR) and various summer
institutes. The white paper could Include a call for a comprehensive assessment
of these techniques. We need to think carefully about the right venue for such a
white paper, where would it have the most impact. The subgroup believes it would
need to approach editors directly to ascertain this.

>>>>

Group discussion started with a position statement by each person surrounding
what they had learned from their software development experiences and how those
lessons might be translated into actionable outcomes. Participants came from a
range of backgrounds, and represented multiple software and training
initiatives, including Software Carpentry (SwC), Data Carpentry (DC), Open Science for
Synthesis (OSS), the Community Surface Dynamics Modeling Systems group, ROpenSci,
DataONE, the HDF Group, and others. Some of these software experiences were
focused on development of new products for use in the sciences (e.g., ROpenSci,
CSDMS), and these recounted the difficulties of engaging with disciplinary
scientists in writing software. Software was clearly utilized broadly across the
various science disciplines represented, and it was developed within those
disciplines as well. Some researchers created software for statistical analysis
and modeling, while others used it to control instrumentation, collate data
across networks, collect information from respondents, and many other uses.
There was broad consensus that, within the disciplines represented, formal
training in any type of software development or engineering was rare among the
practitioners; most are self-taught, and develop software to get another job
done. Any ancillary utility of the software outside of the specific science
target was generally unplanned and few researchers would want to invest more
time to make their own software more re-usable.

@@

Three main actions were identified by Subgroup A that would be of interest to
participants and benefit the research community. These focused on the
desire to amplify the current community efforts in training and software
engineering by connecting the current training initiatives (e.g., SwC, DC, 
rOpenSci, OSS) that are of different durations at appropriate stages.
Generally it was felt that a modicum of interaction between short term workshops
(SwC, DC), medium term trainings (e.g., OSS), and longer-term courses (e.g.,
BIDS Data Science) would benefit from coordinating curriculum, discussing and
aligning prerequisites, and coordinating timing of courses. A training
coordination effort would go a long way towards amplifying the value of the
individual efforts and make them all more effective.

@@

Subgroup B then focused on actions it could take. It discussed development of a
white paper that describes a matrix approach to training (multi-day, multi-week,
semester). The white paper might include a survey of existing techniques. There
are many, some dating back many years, for example the Interuniversity
Consortium for Political and Social Research (ICPSR) and various summer
institutes. The white paper could include a call for a comprehensive assessment
of these techniques. We need to think carefully about the right venue for such a
white paper, where would it have the most impact. The subgroup believes it would
need to approach editors directly to ascertain this.
